Qualitative Probabilistic Programming

### 2015-02-25 paper/paper.tex
more paper

  \begin{verbatim}
  PointType = Vector(2, Double)
  ClusterType = Categorical(3)

  Notice that we declared two types (\texttt{PointType} and \texttt{ClusterType})
  and two random functions (\texttt{get\_cluster} and \texttt{get\_point}).
  Type annotations are necessary for random functions.  The type
  \texttt{Vector(2, Double)} expresses the fact that the points are 2-dimensional,
  and the type \texttt{Categorical(3)} expresses the fact that there are 3
  possible clusters (so a cluster may be either 0, 1, or 2).  We do not know
  the distribution over clusters or over points given a cluster.
  We will fill in the \texttt{get\_cluster} function with
  a learned categorical distribution, and we will fill in the
  \texttt{get\_point} function with a learned unknown linear function with noise.

@@

    For unknown functions, the class of functions used is a kind
    of generalized linear model.  We assume that the result
    is distributed from some exponential family.  We label
    some subset of the sufficient statistics as \emph{features}.  The natural
    parameters corresponding to non-features are constant, while natural
    parameters corresponding to features are determined as an affine
    function of the features of the arguments.

@@

@@

get_point = randFunction(ComponentsType, PointType)

@@

  \subsection{Hidden Markov model}
\begin{verbatim}
num_states = 5
num_observations = 3
sequence_length = 10

@@

\begin{verbatim}
n_classes = 10
n_words = 10000
n_words_per_document = 1000

@@

>>>>

  \begin{verbatim}
  dim = 2
  nclusters = 3
  PointType = Vector(dim, Double)
  ClusterType = Categorical(nclusters)

  get_point = rand_function(ClusterType, PointType)

  def sample():
    cluster = uniform_categorical(nclusters)
    return (cluster, get_point(cluster))
  \end{verbatim}

  Notice that we declared two types (\texttt{PointType} and \texttt{ClusterType})
  and one random function \texttt{get\_point}).
  Type annotations are necessary for random functions.  The type
  \texttt{Vector(2, Double)} expresses the fact that the points are 2-dimensional,
  and the type \texttt{Categorical(3)} expresses the fact that there are 3
  possible clusters (so a cluster may be either 0, 1, or 2).  We assume that
  the disttribution over clusters is uniform, but we do not know the distribution
  of points in each cluster.
  We will fill in the
  \texttt{get\_point} function with a learned unknown linear function with noise.

@@

    For unknown functions, the class of random functions used is a kind
    of generalized linear model.  We assume that the result
    is distributed from some exponential family whose natural
    parameters are determined from the arguments.  We label
    some subset of the sufficient statistics as \emph{features}.  The natural
    parameters corresponding to non-features are constant, while natural
    parameters corresponding to features are determined as an affine
    function of the features of the arguments.

@@

We randomly generated parameters for this example 100 times, and each time took 100 samples and then ran 10 EM iterations.  The accuracy is defined as the maximum percentage of points assigned to the correct cluster, for any permutation of clusters.  On average, accuracy increased in each EM iteration, as shown it this graph:

\begin{center}
\includegraphics[scale=0.5]{cluster_accuracy.png}
\end{center}

\subsection{Naive Bayes}

\begin{verbatim}
nfeatures = 10
FeaturesType = Vector(nfeatures, Bool)

get_class = rand_function(Unit, Bool)

class_1_features = rand_function(Unit, FeaturesType)
class_2_features = rand_function(Unit, FeaturesType)

def sample():
  which_class = get_class()
  if which_class:
    features = class_1_features()
  else:
    features = class_2_features()
  return (which_class, features)
\end{verbatim}

The naive Bayes model is similar to the clustering model.  We have two classes and a feature
distribution for each.  Since each feature is boolean, we will learn
a different categorical distribution for each class.

(figure should show average classification accuracy)

@@

get_point = rand_function(ComponentsType, PointType)

@@

(figure should show accuracy by EM iteration.  Accuracy can be measured by distances between different factor coefficients, when ordered according to some permutation)
  \subsection{Hidden Markov model}
\begin{verbatim}
num_states = 5
num_observations = 3
sequence_length = 10

@@

\begin{verbatim}
n_classes = 10
n_words = 100
n_words_per_document = 1000

@@

In this example, we use the unknown function \texttt{class\_to\_word} to map classes to word distributions.  Note that each column of the matrix of parameters for \texttt{class\_to\_word} will represent a categorical distribution over words, and there will be one column for each class.

(figure should show accuracy over time.  Accuracy can be measured as distance between the learned categorical distributions, for some permutation of classes)

\subsection{Neural Network}

\begin{verbatim}
input_dim = 100
hidden_dim = 20
output_dim = 2

InputType = Categorical(input_dim)
HiddenType = Categorical(hidden_dim)
OutputType = Categorical(output_dim)

input_to_hidden = rand_function(InputType, HiddenType)
hidden_to_output = rand_function(HiddenType, OutputType)

def sample(input_layer):
  hidden_layer = input_to_hidden(input_layer)
  output_layer = hidden_to_output(hidden_layer)
  return ((), output_layer)
\end{verbatim}

\subsection{A more complex model}

TODO: come up with a more complex model.  Show how this model can be iteratively improved by adding information.  It would make sense for this to be some sort of data science type
